---
title: "ETC5513_Tutorial2_v2"
author: "Waranya Ratirotjanakul"
format: html
editor: visual
---

# Reading Avian Influenza Data

```{r}
#| echo: true
#| warning: false
#| message: false

library(tidyverse)
dat <- read_csv("~/Monash_R/ETC5513_Workshop/ETC5513_WS2/WS2_data/avian_influenza_numbers.csv")
```

```{r Reading data}
# return the first n number of row in dataset
head(dat, 10) # display only 10 rows
```

```{r}
# provides us a snapshot of the data 
glimpse(dat) # get the information or dataset, e.g., 
```

```{r}
# quick way for us to check the variable types
str(dat)
```

# R inline command
The data has `r ncol(dat)` variables and `r nrow(dat)` rows.

`r (code)`

### Why is it important to know the dimension of your dataset?
To determine the complexity og your model, validate your result and for exploratory data analysis purposes.

### What are the variable names in the dataset?
There are 26 variables in total.

```{r}
names(dat)

dat$Month
dat$Australia
```
- **Australia:** The number of Inflienza count recorded in Australia for the particular month and year.
- **Month:** The month and year that hte observation is recorded in mm/dd/yy format.

# Exercise 4: Data Wrangling

```{r}
library(tidyverse)
data_cleaned <- dat |>
                  select(Month, Australia, Egypt, `United States`)
```

```{r}
str(data_cleaned)
```
There are four variables which Month is in character, while the others are numeric.
- **Month:** `r class(dat$Month)`
- **Australia**
- **Egypt**
- **United States**

Do you think the variable attributes are correct?
- No, the month should be in date format.

```{r}
library(lubridate)
data_cleaned <- dat |>
                  select(Month, Australia, Egypt, `United States`) |>
                  mutate(Month = mdy(Month)) |>
                  filter(!is.na(Month))
data_cleaned
```

```{r}
library(kableExtra)
data_cleaned |>
  summarise(Australia = sum(Australia),
            Egypt = sum(Egypt),
            `United States` = sum(`United States`)) |>
  kable(caption = "Summary of number of cases of Avian Influenza") |>
  kable_styling()
```

```{r}
library(ggplot2)
data_cleaned |>
  ggplot(aes(x = Egypt)) +
  geom_histogram(binwidth = 5)

data_cleaned |>
  pivot_longer(-Month) |>
  ggplot(aes(x = value, fill = name)) +
  geom_histogram(binwidth = 5, position = 'dodge') +
  facet_wrap(~name)

data_cleaned |>
  pivot_longer(-Month) |>
  ggplot(aes(x = Month, y = value, fill = name)) +
  geom_col()
  facet_wrap(~name)
```

